### 2013-03-28 doc/sosp13/intro.tex
edits

\vspace{0.5ex}
\noindent{\bf High network communication latency.} The performance challenges
of frequent communication is exacerbated by the network latency relative to
processor performance. Latency of commodity networks runs anywhere from one to
hundreds of microseconds, which is much more than superscalar execution with
sophisticated memory hierarchies can tolerate. Since irregular application
tasks encounter remote references dynamically during execution and must
resolve them before making further progress, stalls are frequent and lead
severely underutilized compute resources.

While some irregular applications can be restructured to better exploit
locality, aggregate requests to increase network message size, and manage the
additional challenges of load balance and synchronization, the effort required
to do so is formidable and involves knowledge and skills pertaining to
distributed systems far beyond those of most application programmers. Luckily,
many of the important irregular applications (e.g., graph processing, our
focus in this paper) naturally offer large amounts of concurrency. This
immediately suggests taking advantage of concurrency to tolerate the latency
of data movement by overlapping computation with communication.

@@

\Grappa leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system and an
off-the-shelf user-mode InfiniBand device driver stack~\cite{OFED}. MPI is
used for process setup and tear down. GASNet~\cite{gasnet} is used as the
underlying mechanism for remote memory reads and writes using active message
invocations. To this commodity hardware and software mix \Grappa adds three
main software components: (1) a \emph{lightweight tasking\/} layer that
supports a context switch in as few as 38ns and distributed global load
balancing; (2) a \emph{distributed shared memory\/} layer that supports normal
access operations such as \emph{read\/} and \emph{write\/} as well as
synchronizing operations such as \emph{fetch-and-add\/}~\cite{fetchandadd};
and (3) a \emph{communication\/} layer that combines short messages to
achieve peak bandwidth on commodity networks. As we will show later, \Grappa
can tolerate latencies way beyond that of the network. Therefore, \Grappa can
afford to \emph{trade latency for throughput}. By {\em increasing\/} latency
in key components of the system we are able to: increase effective random
access memory bandwidth by delaying and aggregating messages; increase
synchronization rate by delegating atomic operations to gatekeeper cores, even
when referencing node-local global data; and improve load balance by
tolerating the delays incurred when work-stealing.

>>>>

\vspace{0.5ex} \noindent{\bf High network communication latency.} The
performance challenges of frequent communication is exacerbated by the network
latency relative to processor performance. Latency of commodity networks runs
anywhere from a few to hundreds of microseconds, which is much more than
superscalar execution (even with sophisticated memory hierarchies) can
tolerate. Since irregular application tasks encounter remote references
dynamically during execution and must resolve them before making further
progress, stalls are frequent and lead to severely underutilized compute
resources.

While some irregular applications can be manually restructured to better
exploit locality, aggregate requests to increase network message size, and
manage the additional challenges of load balance and synchronization, the
effort required to do so is formidable and involves knowledge and skills
pertaining to distributed systems far beyond those of most application
programmers. Luckily, many of the important irregular applications (e.g.,
graph processing, our focus in this paper) naturally offer large amounts of
concurrency. This immediately suggests taking advantage of concurrency to
tolerate the latency of data movement by overlapping computation with
communication.

@@

\Grappa leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system and an
off-the-shelf user-mode InfiniBand device driver stack~\cite{OFED}. MPI is
used for process setup and tear down. GASNet~\cite{gasnet} is used as the
underlying mechanism for remote memory reads and writes using active message
invocations. \Grappa adds three main software components: (1) a
\emph{lightweight tasking\/} layer that supports a context switch in as little
as 38ns and distributed global load balancing; (2) a \emph{distributed shared
memory\/} layer that supports normal access operations such as \emph{read\/}
and \emph{write\/} as well as synchronizing operations such as
\emph{fetch-and-add\/}~\cite{fetchandadd}; and (3) a \emph{communication\/}
layer that combines short messages to achieve peak bandwidth on commodity
networks. As we will show later, \Grappa can tolerate latencies way beyond
that of the network. Therefore, \Grappa can afford to \emph{trade latency for
throughput}. By {\em increasing\/} latency in key components of the system we
are able to: increase effective random access memory bandwidth by delaying and
aggregating messages; increase synchronization rate by delegating atomic
operations to gatekeeper cores, even when referencing node-local global data;
and improve load balance by tolerating the delays incurred when work-stealing.
